Algorithms for Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2014-10-14 paper/paper.tex
Minor edits to paper

This paper is an extension of our earlier conference
publication~\cite{Malawski-SC12}.
Here we present a more detailed performance evaluation, including the discussion
of results based on varying deadline and budget constraints. We also report on new
results and discussion of task granularity in the context of the resource
billing cycle, which is an important aspect of scheduling and resource provisioning
on cloud infrastructures.

@@

Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
Kim2011,Chen2011}) also allow to scale the cloud resource pool or to extend the
capabilities of clusters using cloud-burst techniques. Our approach is different
in that we consider workflows, while policy based approaches typically consider
bags of independent tasks or unpredictable batch workloads. This enables us to
take advantage of workflow-aware heuristics that cannot be applied to
independent tasks.

@@

Our work can also be regarded as an extension of the budget-constrained 
workflow scheduling~\cite{Sakellariou2007}, the cost optimization for workflows 
under deadline constraints~\cite{Yu2005}, 
or methods of stretching and compacting the workflow to optimize time and cost~\cite{Lee2013}, 
in the sense that we are dealing with workflow ensembles
and the deadline constraint is added. 
Interesting discussion of uncertainties and the way to address them using Monte Carlo approach 
is also presented in~\cite{Zheng2013b}, where the objective is makespan optimization.

Multiple workflows has been also considered in previous research, 
either in the context of heterogeneous systems~\cite{Zhao2006} 
where makespan and fairness optimization is applied, 
or in the context of grids~\cite{Hirales2010} 
where multiple heuristics are evaluated with respect to the makespan of schedules they produce. 
Scheduling multiple workflows in~\cite{Duan2012,Duan2014} applies multiobjective optimization 
of time and cost under storage bandwidth and capacity constraints with the use of heuristics based on game theory. 
The approach proposed in~\cite{TolosanaCalasanz20121300} focuses on enforcement of QoS 
measured as throughput of processing by individual workflows. 
The problem of workflow admission addressed in~\cite{Zheng2013a} is similar to the one solved by our algorithms, 
but the difference is that we address the problem from the end-user perspective, 
while in their work they assume that the resource provider decides 
whether to admit or reject the workflow based on the resource estimates and the deadline constraints.

@@

Noteworthy are also methods based on mathematical programming, 
using linear programming or mixed-integer programming. 
Such method is applied to scheduling workflows on hybrid clouds using time discretization~\cite{Genez13b}, 
but the model presented there is limited to small scale workflows. 
Large-scale applications on hybrid clouds are addressed in~\cite{VandenBossche2010,VandenBossche2013973}, 
but their model considers bag-of-task applications instead of workflows. 
The cloud bursting scenario where private cloud is combined with a public one, described in~\cite{BittencourtM11}, 
addresses workflows, and the goal is to minimize the cost while keeping the deadline. 
In our related paper~\cite{optimization-ppam13}, we also use the mixed-integer programming approach 
to schedule multi-level deadline-constrained workflows, but workflow ensembles are not considered.

@@

The analysis of related work is summarized in Table~\ref{tab:related}, where we classify the methods based on 7 categories. 
Infrastructure type can be grid or cloud, where grid includes also utility grids 
and heterogeneous computing systems where cost can be also relevant (such as in~\cite{Sakellariou2007}). 
Application types of interest are bags of tasks, single workflows or multiple workflows. 
The workload can be static, where the scheduling process can plan the tasks in advance, 
or dynamic, where the algorithm has to deal with unpredictable stream of jobs arriving. 
The provisioning methods are divided into dynamic and static approaches, 
where static means that the provisioning plan is prepared prior to execution. 
For simplification, we classify all the algorithms for grid and heterogeneous systems into the static category, 
since these systems do not allow for creation of new resources on demand, 
even if some mechanisms for dynamic provisioning such as pilot jobs are available. 
The related work addresses multiple optimization objectives, 
including time, monetary cost, reliability, energy or fairness, 
and there are examples of multicriteria optimization among them. 
Most of the scheduling approaches consider also constraints for optimization, 
and most commonly addressed ones are budget and deadline, 
whereas such additional constraints as bandwidth, storage capacity or job throughput are also present. 
Finally, we divide the algorithmic methods into heuristics, 
including various list scheduling techniques based on graph analysis, 
metaheuristics that use genetic or evolutionary algorithms, 
and the approaches using mathematical programming such as linear 
or mixed-integer programming for which ready to use solvers exist.

The analysis of the table suggests that there are many approaches to workflow scheduling 
and resource provisioning on clouds, but there is still room for improvement. 
New infrastructure types, including hybrid and federated clouds should be investigated, 
more emphasis should be put on problems with multiple workflows, 
and new methods for combining static and dynamic workloads and algorithms offer potentially interesting area of research. 
Finally, other objectives and constraints in addition to budget and execution time can be explored.

@@

>>>>

This paper is an extension of our earlier conference
publication~\cite{Malawski-SC12}.
Here we present a more detailed performance evaluation, including the discussion
of results based on varying deadline and budget constraints. We also report on new
results and discussion of task granularity in the context of the resource
billing cycle, which is an important aspect of scheduling and resource provisioning
on cloud infrastructures. The related work section and the algorithm examples
have been also extended to provide a better positioning and increase the readability of the paper.

@@

Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
Kim2011,Chen2011}) also enable scaling the cloud resource pool or to extending the
capabilities of clusters using cloud-burst techniques. Our approach is different
in that we consider workflows, while policy based approaches typically consider
bags of independent tasks or unpredictable batch workloads. This enables us to
take advantage of workflow-aware heuristics that cannot be applied to
independent tasks.

@@

Our work can also be regarded as an extension of budget-constrained 
workflow scheduling~\cite{Sakellariou2007}, cost optimization for workflows 
under deadline constraints~\cite{Yu2005}, 
or methods of stretching and compacting the workflow to optimize time and cost~\cite{Lee2013}, 
in the sense that we are dealing with workflow ensembles
and the deadline constraint is added. 
An interesting discussion of uncertainties and the way to address them using a Monte Carlo approach 
is also presented in~\cite{Zheng2013b}, where the objective is makespan optimization.

Scheduling optimization for multiple workflows has been also considered in previous research, 
either in the context of heterogeneous systems~\cite{Zhao2006} 
where makespan and fairness optimization is applied, 
or in the context of grids~\cite{Hirales2010} 
where multiple heuristics are evaluated with respect to the makespan of schedules they produce. 
Scheduling multiple workflows in~\cite{Duan2012,Duan2014} applies multiobjective optimization 
of time and cost under storage bandwidth and capacity constraints with the use of heuristics based on game theory. 
The approach proposed in~\cite{TolosanaCalasanz20121300} focuses on enforcement of QoS 
measured as throughput of processing by individual workflows. 
The problem of workflow admission addressed in~\cite{Zheng2013a} is similar to the one solved by our algorithms, 
but the difference is that we address the problem from the end-user perspective, 
while in their work they assume that the resource provider decides 
whether to admit or reject the workflow based on the resource estimates and the deadline constraints.

@@

Also noteworthy are methods based on mathematical programming, 
using linear programming or mixed-integer programming. 
Such a method is applied to scheduling workflows on hybrid clouds using time discretization in~\cite{Genez13b}, 
but the model presented there is limited to small scale workflows. 
Large-scale applications on hybrid clouds are addressed in~\cite{VandenBossche2010,VandenBossche2013973}, 
but their model considers bag-of-task applications instead of workflows. 
The cloud bursting scenario described in~\cite{BittencourtM11}, where a private cloud is combined with a public one, 
addresses workflows, and the goal is to minimize the cost while satisfying the deadline. 
In our related paper~\cite{optimization-ppam13}, we also use the mixed-integer programming approach 
to schedule multi-level deadline-constrained workflows, but workflow ensembles are not considered.

@@

The analysis of related work is summarized in Table~\ref{tab:related}, where we classify the methods based on 7 categories. 
Infrastructure type can be grid or cloud, where grid includes utility grids 
and heterogeneous computing systems where cost can be relevant (such as in~\cite{Sakellariou2007}). 
Application types of interest are bags of tasks, single workflows or multiple workflows. 
The workload can be static, where the scheduling process can plan the tasks in advance, 
or dynamic, where the algorithm has to deal with unpredictable stream of jobs arriving. 
The provisioning methods are divided into dynamic and static approaches, 
where static means that the provisioning plan is prepared prior to execution. 
To simplify, we classify all the algorithms for grid and heterogeneous systems into the static category, 
since these systems do not allow for creation of new resources on demand, 
even if some mechanisms for dynamic provisioning such as pilot jobs are available. 
The related work addresses multiple optimization objectives, 
including time, monetary cost, reliability, energy or fairness, 
and there are examples of multi-criteria optimization among them. 
Most of the scheduling approaches also consider constraints for optimization, 
and the most commonly addressed ones are budget and deadline. 
Other constraints, such as bandwidth, storage capacity or job throughput, are also present. 
Finally, we divide the algorithmic methods into heuristics, 
including various list scheduling techniques based on graph analysis, 
metaheuristics that use genetic or evolutionary algorithms, 
and the approaches using mathematical programming such as linear 
or mixed-integer programming for which ready to use solvers exist.

An analysis of the table suggests that there are many approaches to workflow scheduling 
and resource provisioning on clouds, but there is still room for improvement. 
New infrastructure types, including hybrid and federated clouds should be investigated, 
more emphasis should be put on problems with multiple workflows, 
and new methods for combining static and dynamic workloads and algorithms offer potentially interesting area of research. 
Finally, other objectives and constraints in addition to budget and execution time can be explored.

@@

The decision to use a simulator instead of actually executing workflows on clouds 
was motivated by our intention to perform an extensive analysis of the algorithms
in various aspects. This results in a very large parameter space. For example, the analysis 
presented in Section~\ref{sec:variances} required 525,000 simulations 
of workflow ensembles, each ensemble consisting of 50 workflows with up to 1000 tasks per workflow.
Not only would this be impractical to run on a real infrastructure for obvious reasons, 
but it would also not enable us to study interesting aspects of the problem, such as the influence 
of task granularity presented in Section~\ref{sec:granularity}. 
Our use of a simulator, although quite compute-intensive itself, enabled us
to study all these aspects in detail.
